---
title: "NST pitch preprocessing"
author: "Nadja Althaus"
date: "14/02/2022"
output: html_document
---

In this file we normalise and smooth pitch contours of disyllabic nouns extracted from the NST corpus.
The aim is to assimilate Nencheva's (2020) treatment of pitch contours.
This involves Tukey's median smoothing, then identifying the 30 'most interesting' points (Nencheva, 2020), which are the points with the largest slope change between consecutive time slots. Note that while Nencheva used 10 such points, we get better results with 30 points for the present task, perhaps because Acc2 contours are more complex.
The pitch contours are interpolated between these 'most interesting' points, and normalized to deal with absolute pitch height differences.



```{r libraries}
library(ggplot2)
library(stringr)
library(stats)
library(zoo)

```

```{r load.data}


NSTData=read.csv("PitchDF.csv", fileEncoding='iso-8859-1',  header=TRUE, stringsAsFactors=FALSE)

```

```{r read.pitch.file}



read_pitch <- function(pitchfilename){
  # read the pitch from an output file of praat f0 extraction script
  # parameters:
  # pitchfilename - .txt file holding pitch info
  
  
  #need to skip 2 lines that hold info
  tryCatch({
  pitchdata=read.csv(pitchfilename,sep=" ", skip=2, col.names = c("t","p"), fill=TRUE, stringsAsFactors=FALSE , header=FALSE)
  
  
  #then break down info from first 2 lines: start/end timestamps in format startt--endt and number of samples
  pfile <- file(pitchfilename,"r")
  start_end <- readLines(pfile,n=1)
  vals=strsplit(start_end, '--')
  
  startt=vals[[1]][1]
  endt=vals[[1]][2]
  nsamples <-readLines(pfile,n=1)
  close(pfile)
  outlist=list(startt,endt, nsamples, pitchdata)
  return(outlist)
  }, error=function(e){return(list())})
}


```


```{r prep.functions}


clip_na <- function(mylist){
  #remove leading and trailing NA slots
  non_na_idx=which(!is.na(mylist))
  removed_leading=mylist[non_na_idx[1]:length(mylist)]
  non_na_idx=which(!is.na(removed_leading))
  removed_trailing=removed_leading[1:non_na_idx[length(non_na_idx)]]
  return(removed_trailing)
}



find_informative_points <- function(pitchlist, n){
  #returns the n points with the largest change in slope between consecutive items

  x2=1:length(pitchlist)
  x1=append(NA, 1:(length(pitchlist)-1))
  y2=pitchlist
  y1=append(NA,pitchlist[1:length(pitchlist)-1])
  diffy=y2-y1
  diffx=x2-x1
  #div will contain pairwise slopes between adjacent points
  div=diffy/diffx
  div2=append(NA,div[1:length(div)-1])#shifted
  diffdiv=abs(div2-div)
  
  #get indices of div sorted by slope, then reverse to get the 10 points with lowest slope
  odiffdiv=rev(order(diffdiv))
  returnlist=sort(odiffdiv[1:n])
  if (length(returnlist)>length(pitchlist)){
    returnlist=returnlist[1:length(pitchlist)]
  }
  return(returnlist)
                          
  
}




```


```{r final.prepFunction}



prepPitch <- function (pitchdata, npoints){
  #puts all prep steps together
  #removes leading/trailing NA
  #followed by smoothing, identifying the n most interesting points, interpolation around these, and 
  #normalisation
  clipped_pitch=clip_na(pitchdata$p)
  nsamp=length(clipped_pitch)
  smooth_pitch=smooth(na.approx(clipped_pitch))
  informative_idx=find_informative_points(smooth_pitch, npoints)
  informative_y=smooth_pitch[informative_idx]
  x_scaled=round(((informative_idx)/nsamp)*30)
  
  y_interpolate=rep(NA,30)
  y_interpolate[x_scaled]=informative_y
  
  if (is.na(y_interpolate[1])){
    y_interpolate[1]=smooth_pitch[1]
  }
  if (is.na(y_interpolate[length(y_interpolate)])){
    y_interpolate[length(y_interpolate)]=smooth_pitch[length(smooth_pitch)]
  }
  y_interpolate=smooth(na.approx(y_interpolate))
  
  mn=min(y_interpolate)
  mx=max(y_interpolate)-mn
  y_interpolate=(y_interpolate-mn)/mx
  

  return(y_interpolate)
}


```

```{r}

#process the entire pitch file corpus
PitchArray=array(dim=c(dim(NSTData)[1],30))
for (i in 1:dim(NSTData)[1]){
  print(i)
  
  pitchfilename=NSTData$PitchFile[i]
  print(pitchfilename)
  pitch=read_pitch(pitchfilename)

  
  pitchdata=pitch[[4]]
  if (sum(is.na(pitchdata))==dim(pitchdata)[1]){
    print('Does not have pitch data')
    next
  }
  smoothpitch <-prepPitch(pitchdata, 20)
  
  PitchArray[i,]=smoothpitch
  
}
```

```{r AddToSvenskData}
for (i in 1:30){
 NSTData[paste0('S',i, sep="")]=PitchArray[,i]
  
}



```


